---
title: "Counterfactuals for paper"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
mean_ci <- function(x) c("mean" = mean(x), quantile(x, c(0.025, 0.975)))

extract_deaths <- function(x, what, from = "2020-01-01", to, traj = "deaths") {
  trajectories <- x[[what]]$trajectories
  s_date <- sircovid::sircovid_date(c(from, to))
  ret <- trajectories$state[traj, , trajectories$date == s_date[2]] -
    trajectories$state[traj, , trajectories$date == s_date[1]]
  mean_ci(ret)
}

extract_mean_fitted_deaths <- function(x, to, from = "2020-01-01",
                                       traj = "deaths") {
  round(extract_deaths(x, "fitted", from, to, traj)[["mean"]])
}

vnapply <- function(X, FUN, ...) {
  vapply(X, FUN, numeric(1), ...)
}

vn3apply <- function(X, FUN, ...) {
  vapply(X, FUN, numeric(3), ...)
}
```

Cumulative deaths England

```{r}
date_end_ld1 <- "2020-08-01"
```


Cumulative fitted deaths in first wave, in England

```{r}
mean_fitted_deaths <- extract_mean_fitted_deaths(agg_counterfactuals,
                                                 to = date_end_ld1)
print(mean_fitted_deaths)
```

Lockdown 1 starting late deaths

```{r}
extract_deaths(agg_counterfactuals, "lockdown_later", to = date_end_ld1)
```

Lockdown 1 starting early deaths

```{r}
extract_deaths(agg_counterfactuals, "lockdown_earlier", to = date_end_ld1)
```

Impact of lockdown 1 starting later, by region

```{r}
tmp <- vn3apply(counterfactuals_by_region, extract_deaths,
                "lockdown_later", to = date_end_ld1)

deaths_region <- vnapply(counterfactuals_by_region,
                         extract_mean_fitted_deaths, to = date_end_ld1)
round((t(tmp) / deaths_region - 1) * 100, 1)
vn3apply(counterfactuals_by_region, extract_deaths, "fitted",
         to = date_end_ld1)
```

Fitted deaths in the second wave, used to compare against counterfactuals

```{r}
date_end <- "2020-12-01"
mean_fitted_deaths <- extract_mean_fitted_deaths(agg_counterfactuals,
                                                 from = date_end_ld1,
                                                 to = date_end)
```

Relax earlier

```{r}
extract_deaths(agg_counterfactuals, "open_earlier",
               from = date_end_ld1, to = date_end) - mean_fitted_deaths
```

Relax later

```{r}
extract_deaths(agg_counterfactuals, "open_later",
               from = date_end_ld1, to = date_end) - mean_fitted_deaths
```

# Carehome contact change % reduction in carehome deaths

```{r}
v <- sprintf("chr_contact_reduce_%s", c(0.25, 0.5, 0.75, 1))
ccr <- vn3apply(v, extract_deaths, x = agg_counterfactuals,
                from = "2020-01-01", to = date_end_ld1,
                traj = "deaths_carehomes")
mean_fitted_carehome_deaths <-
  extract_mean_fitted_deaths(agg_counterfactuals,
                             from = "2020-01-01",
                             to = date_end_ld1,
                             traj = "deaths_carehomes")
impact_ccr <- t(ccr) / mean_fitted_carehome_deaths - 1
round(impact_ccr * 100, 1)
```

# Carehome contact change increase - no impact here

```{r}
v <- sprintf("chr_contact_increase_%s", c(0.25, 0.5, 0.75, 1))
cci <- vn3apply(v, extract_deaths, x = agg_counterfactuals,
                from = "2020-01-01", to = date_end_ld1,
                traj = "deaths_carehomes")
t(cci) - mean_fitted_carehome_deaths
```

# Discussion

How much more effective did contact reduction need to be?

```{r}
apply(impact_ccr, 2, function(x)
  approx(x, c(0.25, 0.5, 0.75, 1), xout = -0.5)$y)
```
