---
title: "Numbers for the paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

format_num <- function(x, d) {
  x <- round(x / d, 1) * d
  prettyNum(x, ",")
}

mean_ci <- function(x) c(mean = mean(x), quantile(x, c(0.025, 0.975)))

extract_value <- function(what, date, sample) {
  sircovid_date <- sircovid::sircovid_date(date)
  sample$trajectories$state[what, , sample$trajectories$date == sircovid_date]
}

vnapply <- function(X, FUN, ...) {
  vapply(X, FUN, numeric(1), ...)
}

paper_date <- "2020-12-02"
start_data <- "2020-03-16"
end_first_wave <- "2020-08-01"
# dates are midpoint of first and second lockdown
# lockdown 1: 2021-03-23 to 2020-05-11
# lockdown 2: 2020-11-05 -> paper date = "2020-12-02"
ld_dates <- data.frame(name  = c("ld1", "ld2"),
                       start = sircovid_date(c("2020-03-23", "2020-11-05")),
                       end   = sircovid_date(c("2020-05-11", paper_date)))
ld_dates$mid <-
  as.character(sircovid_date_as_date((ld_dates$start + ld_dates$end) / 2))

pop <- sapply(names(agg_samples),
              function(x) sum(sircovid:::sircovid_population(x)))
```

# Abstract

% of infected admitted to hospital
```{r}
infected <- extract_value("infections", paper_date, agg_samples$england)
admitted <- extract_value("admitted", paper_date, agg_samples$england) +
  extract_value("diagnoses", paper_date, agg_samples$england)

round(mean_ci(admitted / infected) * 100, 2)
```

% of infected that died

```{r}
died <- extract_value("deaths", paper_date, agg_samples$england)
round(mean_ci(died / infected) * 100, 2)
```

% pop infected to 1st August

```{r}
infected_firstwave <- extract_value("infections", end_first_wave,
                                    agg_samples$england)

round(mean_ci(infected_firstwave / pop["england"]) * 100, 2)
```

% infected in second wave

```{r}
infected_secondwave <- infected - infected_firstwave

round(mean_ci(infected_secondwave / pop["england"]) * 100, 2)
```

Total % infected (to 2nd December 2020)

```{r}
round(mean_ci(infected / pop["england"]) * 100, 2)
```


# Introduction

Deaths by region from before RTM time series starts

```{r}
early_deaths <- c("london" = 37,
                  "east_of_england" = 12,
                  "midlands" = 20,
                  "north_east_and_yorkshire" = 3,
                  "north_west" = 11,
                  "south_east" = 13,
                  "south_west" = 3)
```


```{r}
tot_deaths <- sum(agg_data$england$full$deaths, na.rm = TRUE) +
                sum(early_deaths)
tot_deaths
```
```{r}
tot_deaths / sum(sircovid:::sircovid_population("england")) * 1e5
```

# Results

## 1. Epidemic trajectory

R0 in England

```{r}
extract_Rt_est <- function(date, Rt_output, what) {
  ret <- Rt_output[[what]][Rt_output$date == sircovid::sircovid_date(date)]
  round(mean_ci(ret), 2)
}

extract_Rt_est(start_data, agg_Rt_outputs$england, "eff_Rt_all")
```

Regional epidemic start dates

```{r}
start_dates <- sapply(samples[-8], function(x) {
  start_date <- mean_ci(x$pars[, "start_date"])
  as.character(sircovid::sircovid_date_as_date(start_date))
  })

t(start_dates)
```


```{r}
start_date <- lapply(agg_samples, function(x) x$pars[, "start_date"])

mean_ci(start_date$midlands - start_date$london)
```


Death toll per 100,000 during the first wave (up to 1st August 2020)

```{r echo = FALSE}
first_wave_deaths <- sapply(agg_samples, function(x) {
  mean_ci(extract_value("deaths", end_first_wave, x))
})

round(t(first_wave_deaths) / pop * 1e5, 1)
```


Proportion of London population in care homes compared to national average

```{r}
i_eng <- which(names(agg_samples) == "england")
chr_pop <- sapply(agg_samples[-i_eng], function(x) {
  ret <- x$predict$transform(x$pars[1, ])
  ret$carehome_residents
})

chr_pop <- c(chr_pop, "england" = sum(chr_pop))

round(chr_pop / pop * 1e5, 0)
```

Proportion of first wave deaths that occurred in care homes (England)

```{r}
prop_chr_deaths <- extract_value("deaths_carehomes", end_first_wave,
                                 agg_samples$england) /
  extract_value("deaths", end_first_wave, agg_samples$england)

paste0(round(mean_ci(prop_chr_deaths), 3) * 100, "%")

i <- as.Date(agg_data$england$full$date_string) <= as.Date(end_first_wave)
sum(agg_data$england$full$deaths_carehomes[i], na.rm = TRUE) /
  sum(agg_data$england$full$deaths[i], na.rm = TRUE)
```


Timing of first-wave peak in care homes relative to hospital deaths

```{r}
calc_peak <- function(sample, what, end_date) {
  i_date <- sample$trajectories$date <= sircovid::sircovid_date(end_date)
  daily_deaths <- sample$trajectories$state[what, , i_date]
  dates <- sample$trajectories$date[i_date]
  dates[apply(daily_deaths, 1, which.max)]
}

peak_diff <- function(sample, end_date) {
  chw <- calc_peak(sample, "deaths_carehomes_inc", end_first_wave)
  hosp <- calc_peak(sample, "deaths_hosp_inc", end_first_wave)
  mean_ci(chw - hosp)
}

t(sapply(agg_samples, peak_diff, end_first_wave))
```

Risk of infection in care homes

```{r}
calc_infections_by_age <- function(sample, S_i) {
  S <- sample$trajectories$state[S_i, , ]
  infections <- S[, 1] - t(S)
  infections_inc <- t(apply(infections, 2, diff))
  infections_inc[, 1] <- 0
  risk <- infections_inc / S[, -ncol(S)]
  list(infections_inc = cbind(0, infections_inc),
       risk = risk)
}

calc_S_gen <- function(sample) {
  state <- sample$trajectories$state
  S_names <- grep("^S_", rownames(state), value = TRUE)
  S_names <- setdiff(S_names, "S_CHR")
  apply(state[S_names, , ], c(2, 3), sum)
}

calc_infections_nonCHR <- function(sample) {
  infections_inc_CHR <- calc_infections_by_age(sample, "S_CHR")$infections_inc
  infections_inc <- sample$trajectories$state["infections_inc", , ]
  infections_inc - infections_inc_CHR
}

calc_risk_inf_gen <- function(sample) {
  infections_inc_gen <- calc_infections_nonCHR(sample)
  S_gen <- calc_S_gen(sample)
  infections_inc_gen[, -1] / S_gen[, -ncol(S_gen)]
}

peak_risk_date <- function(sample, end_date) {
  i_dates <- sample$trajectories$date <= sircovid::sircovid_date(end_date)
  sample$trajectories$date <- sample$trajectories$date[i_dates]
  sample$trajectories$state <- sample$trajectories$state[, , i_dates]

  gen <- calc_risk_inf_gen(sample)
  chr <- calc_infections_by_age(sample, "S_CHR")$risk

  extract_peak <- function(x) {
    ret <- mean_ci(sample$trajectories$date[apply(x, 1, which.max)])
    ret <- sircovid::sircovid_date_as_date(ret)
    as.character(ret)
  }

  gen_peak <- extract_peak(gen)
  names(gen_peak) <- sprintf("gen_%s", names(gen_peak))
  chr_peak <- extract_peak(chr)
  names(chr_peak) <- sprintf("chr_%s", names(chr_peak))
  c(gen_peak, chr_peak)
}

peak_risk <- sapply(agg_samples, peak_risk_date, end_first_wave)
t(peak_risk)
```


$R_t ^ {eff}$ in first lockdown compared to second

```{r}
ld_dates$mid

lapply(ld_dates$mid, extract_Rt_est, agg_Rt_outputs$england, "eff_Rt_general")
```

$R_t$ in second lockdown

```{r}
lapply(ld_dates$mid, extract_Rt_est, agg_Rt_outputs$england, "Rt_general")
```

## 2. Severity and hopitalisation

Regional infection fatality ratio (IFR) / hospitalisation ratio (IHR) in first wave

```{r}
extract_severity <- function(ifr, what, end_date) {
  mean_ci(ifr[[what]][ifr$date == sircovid::sircovid_date(end_date), ])
}

regional_ifr <- t(sapply(ifr_t, extract_severity, "IFR_AR_all", start_data))

regional_ihr <- t(sapply(ifr_t, extract_severity, "IHR_AR_all", start_data))

list(ifr = round(regional_ifr, 2),
     ihr = round(regional_ihr, 2))
```

How much did hospital mortality improve?

```{r}
reduction_in_severity <- function(ifr, date1, date2) {
  ret <- 1 - (ifr$IFR_AR_all[ifr$date == sircovid::sircovid_date(date2), ] /
    ifr$IFR_AR_all[ifr$date == sircovid::sircovid_date(date1), ])
  mean_ci(ret)
}

round(t(sapply(ifr_t, reduction_in_severity,
               start_data, end_first_wave)) * 100, 1)
```

First wave IFR

```{r}
round(extract_severity(ifr_t$england, "IFR_AR_all", start_data), 2)

round(extract_severity(ifr_t$england, "IFR_AR_all", end_first_wave), 2)
```

Second wave IFR

```{r}
round(extract_severity(ifr_t$england, "IFR_AR_all", paper_date), 2)
```

IFR in care home residents (CHR)

```{r}
round(t(sapply(ifr_t, extract_severity, "IFR_CHR", end_first_wave)), 1)
```

IFR in oldest age group (80+)

```{r}
round(t(sapply(ifr_t, extract_severity, "IFR_80", end_first_wave)), 1)
```

IHR in CHR

```{r}
round(t(sapply(ifr_t, extract_severity, "IHR_CHR", end_first_wave)), 1)
```

IHR in oldest age group (80+)

```{r}
round(t(sapply(ifr_t, extract_severity, "IHR_80", end_first_wave)), 1)
```

# 3. Epidemic size

Total proportion ever infected by region

```{r}
total_infs <- sapply(agg_samples, function(x)
  mean_ci(extract_value("infections", paper_date, x)))

round(t(total_infs) / pop * 100, 1)
```
Population ever infected in CHR by region

```{r}
total_infs_chr <- chr_pop + t(sapply(agg_samples, function(x)
  mean_ci(-extract_value("S_CHR", paper_date, x))))

round(total_infs_chr / chr_pop * 100, 1)
```

Population over 80 ever infected by region

```{r}
i_eng <- which(names(agg_samples) == "england")
pop_80 <- sapply(agg_samples[-i_eng], function(x) {
  ret <- x$predict$transform(x$pars[1, ])
  ret$N_tot[ret$n_age_groups]
})

pop_80 <- c(pop_80, "england" = sum(pop_80))

total_infs_80 <- pop_80 + t(sapply(agg_samples, function(x)
  mean_ci(-extract_value("S_80", paper_date, x))))

round(total_infs_80 / pop_80 * 100, 1)
```

Serology parameters:
Time to seropositivity (days)

```{r}
pars <- samples$london$predict$transform(samples$london$pars[1, ])
(1 / pars$gamma_sero_pre_1 + 1 / pars$gamma_sero_pre_2) / 2
```

Probability of seroconversion, sensitivity / specificity of serological tests

```{r}
pars$p_sero_pos[1]
pars$sero_sensitivity
pars$sero_specificity
```

Daily deaths peak

```{r}
max(agg_data$england$full$deaths, na.rm = TRUE)
```

See Counterfactuals task for the remaining Results numbers

# Discussion

Regional IFR at end of epidemic

```{r}
round(t(sapply(ifr_t, extract_severity, what = "IFR_AR_all", paper_date)), 2)
```

Regional IHR at end of epidemic

```{r}
round(t(sapply(ifr_t, extract_severity, what = "IHR_AR_all", paper_date)), 2)
```

# Supplement

Table S6: Fitted model parameters

Regional epidemic start date

```{r}
extract_params1 <- function(sample, what, digits = 2) {
  tmp <- round(apply(sample$pars[, what, drop = FALSE], 2, mean_ci), digits)
  tmp <- as.data.frame(t(tmp))
  sprintf("%s (%s, %s)", tmp$mean, tmp$`2.5%`, tmp$`97.5%`)
}

extract_params <- function(samples, what, digits) {
  tmp <- sapply(samples[region_names], extract_params1, what = what,
                digits = digits)
  if (!is.null(nrow(tmp))) {
    rownames(tmp) <- what
  }
  tmp
}


sapply(samples[region_names], function(x) {
  tmp <- sircovid::sircovid_date_as_date(mean_ci(x$pars[, "start_date"]))
  as.character(tmp)
})
```

Regional betas and eps

```{r}
extract_params(samples, c(sprintf("beta%s", seq_len(12)), "eps"), 2)
```

Care home parameters

```{r}
extract_params(samples, what = c("m_CHR", "m_CHW"), digits = 7)
```

Transistion probabilities

```{r}
v <- c("H", "H_CHR", "G_D", "G_D_CHR", "ICU", "H_D", "ICU_D",
       "W_D")
extract_params(samples, what = sprintf("p_%s", v), digits = 2)
```

mus

```{r}
extract_params(samples, what = sprintf("mu_%s", c("D", "ICU")),
               digits = 2)
```

The rest

```{r}
extract_params(samples, what = c("p_NC", "rho_pillar2_tests",
                                 "alpha_D", "alpha_H"),
               digits = 4)
```

Table S9: England IFR / IHR by age / CHR

```{r}
format_ifr <- function(what) {
  groups <- c(seq(0, 80, 5), "CHR")
  tmp <- t(sapply(sprintf("%s_%s", what, groups), extract_severity,
           ifr = ifr_t$england, end_date = paper_date))
  tmp <- round(tmp, 3)
  paste0(tmp[, "mean"], "% (", tmp[, "2.5%"], "%, ", tmp[, "97.5%"], "%)")
}

cbind(IHR = format_ifr("IHR"), IFR = format_ifr("IFR"))
```

EpiEstim section:

```{r}
mean_days_infection_to_death
mean_gt_distr
std_gt_distr
```
